---
title: "FUSE Example: Centrality Score Estimation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction

This vignette demonstrates the usage of the FUSE (Fused Unified centrality Score Estimation) package in R. FUSE combines global data centrality learning and local denoising score matching through a shared encoder with two scalar heads.

## Installation

First, install the required packages:

```{r install, eval=FALSE}
# Install torch and other dependencies
install.packages(c("torch", "ggplot2", "umap", "MASS", "patchwork"))
torch::install_torch()

devtools::install_github("vuminhducvmd/FUSE-R")
```

Load the packages:

```{r load}
library(torch)
library(FUSE)
library(ggplot2)
library(umap)
library(MASS)
library(patchwork)
```

## Data Generation

Let's generate synthetic data from a mixture of normal distributions:

```{r data_gen}
# Function to generate mixture of normal data
generate_mixture_normal_data <- function(n = 1000, d = 2, k = 3, test_size = 0.1, cov = NULL, random_state = 42) {
  set.seed(random_state)
  weights <- rep(1/k, k)  # Simple equal weights
  means <- matrix(rnorm(k * d, 0, 5), k, d)

  if (is.null(cov)) {
    covs <- replicate(k, diag(d), simplify = FALSE)  # Identity covariance
  } else if (is.list(cov)) {
    covs <- cov
  } else {
    covs <- replicate(k, cov, simplify = FALSE)
  }

  labels <- sample(1:k, n, replace = TRUE, prob = weights)
  data <- matrix(0, n, d)
  for (comp in 1:k) {
    idx <- which(labels == comp)
    if (length(idx) > 0) {
      data[idx, ] <- mvrnorm(length(idx), means[comp, ], covs[[comp]])
    }
  }

  if (test_size == 0) {
    return(list(data = data, labels = labels))
  }

  train_idx <- sample(1:n, round(n * (1 - test_size)))
  test_idx <- setdiff(1:n, train_idx)

  list(
    train = data[train_idx, ],
    test = data[test_idx, ],
    y_train = labels[train_idx],
    y_test = labels[test_idx]
  )
}

# Generate training data
set.seed(123)
data <- generate_mixture_normal_data(n = 5000, d = 2, k = 4, random_state = 123, test_size = 0)
trainset <- data$data
y_train <- data$labels

# Visualize the data
df_data <- data.frame(x = trainset[, 1], y = trainset[, 2], label = as.factor(y_train))
ggplot(df_data, aes(x = x, y = y, color = label)) +
  geom_point(alpha = 0.6) +
  labs(title = "Synthetic Mixture of Normal Data", x = "X1", y = "X2") +
  theme_minimal()
```

## Model Training

Set up the device and create the FUSE model:

```{r model_setup}
# Set device
device <- if (cuda_is_available()) torch_device("cuda") else torch_device("cpu")
cat("Using device:", device$type, "\n")

# Convert data to torch tensors
train_tensor <- torch_tensor(trainset, dtype = torch_float(), device = device)

# Compute dissimilarity matrix (Euclidean distance)
Dtrain <- torch_cdist(train_tensor, train_tensor, p = 2)

# Create FUSE model
model <- FUSE(input_dim = 2, hidden = c(32, 32), dropout = 0.0)
model$to(device)
```

Train the model:

```{r training}
# Fit the model
history <- model$fit(
  X = train_tensor,
  dissimilarity_matrix = Dtrain,
  partitioned = TRUE,
  allPairs = FALSE,
  num_epochs = 20,
  dsm_resampling = 8,
  anchors_per_pair_global = 64,
  lr = 1e-3,
  device = device,
  quiet = FALSE,
  seed = 379
)

# Plot training history
history_df <- do.call(rbind, lapply(1:length(history), function(i) {
  data.frame(epoch = i, loss_global = history[[i]]$loss_global,
             loss_gradl = history[[i]]$loss_gradl,
             loss_total = history[[i]]$loss_total)
}))

ggplot(history_df, aes(x = epoch)) +
  geom_line(aes(y = loss_global, color = "Global Loss")) +
  geom_line(aes(y = loss_gradl, color = "Local Loss")) +
  geom_line(aes(y = loss_total, color = "Total Loss")) +
  labs(title = "Training History", x = "Epoch", y = "Loss") +
  theme_minimal()
```

## Inference and Visualization

Perform inference at different t values (interpolation between global and local centrality):

```{r inference}
# Inference at different t values
t_values <- c(0.0, 0.25, 0.5, 0.75, 1.0)
scores_list <- lapply(t_values, function(t) {
  scores <- model$inference(train_tensor, t)
  as.numeric(scores$cpu())
})

# Create data frame for plotting
inference_df <- data.frame(
  x = rep(trainset[, 1], length(t_values)),
  y = rep(trainset[, 2], length(t_values)),
  t = rep(t_values, each = nrow(trainset)),
  score = unlist(scores_list)
)

ggplot(inference_df, aes(x = x, y = y, color = score)) +
  geom_point(alpha = 0.6) +
  facet_wrap(~ t, labeller = labeller(t = function(x) paste("t =", x))) +
  scale_color_viridis_c() +
  labs(title = "Centrality Scores at Different t Values",
       x = "X1", y = "X2", color = "Score") +
  theme_minimal()
```

## Contour Plots

Create contour plots showing the homotopy centrality function:

```{r contours, fig.width=12, fig.height=4}
# Plot homotopy contours
plots <- plot_homotopy_contours(model, trainset, t_values = t_values,
                                title = "Homotopy Centrality Contours")

if (is.list(plots)) {
  combined_plot <- wrap_plots(plots) + plot_annotation(title = "Homotopy Centrality Contours - Mixture of Normal Distribution")
  print(combined_plot)
} else {
  print(plots)
}
```

## UMAP Visualization

Visualize the centrality in UMAP space:

```{r umap, fig.width=12, fig.height=8}
# Plot UMAP
umap_plots <- plot_homotopy_umap(model, trainset, y = y_train,
                                 t_values = c(0.0, 0.33, 0.67, 1.0),
                                 title = "UMAP — Labels + Homotopy Centrality per t")

if (is.list(umap_plots)) {
  combined_umap <- wrap_plots(umap_plots) + plot_annotation(title = "UMAP Visualization")
  print(combined_umap)
}
```

## Conclusion

This example demonstrates how to use FUSE for centrality estimation on synthetic data. The model learns both global and local centrality measures and interpolates between them using the homotopy parameter t.

- t = 0: Pure global centrality
- t = 1: Pure local centrality
- 0 < t < 1: Hybrid centrality

The contour plots show how the centrality landscape changes with t, and the UMAP visualizations help understand the multi-scale nature of the learned centrality.