---
title: "FUSE: Fused Unified Centrality Score Estimation"
author: "R Implementation Example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Installation

Install the FUSE package from GitHub:

```{r install, eval=FALSE}
# Install required packages first
install.packages(c("torch", "MASS", "devtools"))
torch::install_torch()

# Install FUSE from GitHub
devtools::install_github("vuminhducvmd/FUSE-R")
```

## Load packages

```{r load}
library(torch)
library(FUSE)
library(MASS)
```

## 2D Mixture of Normal data

### Data generation function

```{r data_gen_function}
generate_mixture_normal_data <- function(n = 1000, d = 2, k = 3, test_size = 0.1, 
                                         cov = NULL, random_state = 42) {
  set.seed(random_state)
  
  weights <- rep(1/k, k)
  means <- matrix(rnorm(k * d, mean = 0, sd = 5), k, d)
  
  if (is.null(cov)) {
    covs <- replicate(k, diag(d), simplify = FALSE)
  } else if (is.list(cov)) {
    covs <- cov
  } else {
    covs <- replicate(k, cov, simplify = FALSE)
  }
  
  labels <- sample(1:k, size = n, replace = TRUE, prob = weights)
  data <- matrix(0, n, d)
  
  for (comp in 1:k) {
    idx <- which(labels == comp)
    if (length(idx) > 0) {
      data[idx, ] <- mvrnorm(n = length(idx), mu = means[comp, ], Sigma = covs[[comp]])
    }
  }
  
  if (test_size == 0) {
    return(list(data = data, labels = labels))
  }
  
  train_idx <- sample(1:n, size = round(n * (1 - test_size)))
  test_idx <- setdiff(1:n, train_idx)
  
  list(
    train = data[train_idx, ],
    test = data[test_idx, ],
    y_train = labels[train_idx],
    y_test = labels[test_idx]
  )
}
```

### Setup device

```{r device}
device <- if (cuda_is_available()) torch_device("cuda") else torch_device("cpu")
cat("Using device:", device$type, "\n")
```

### Generate training data

```{r generate_data}
data_result <- generate_mixture_normal_data(n = 5000, d = 2, k = 4, random_state = 123, test_size = 0)
trainset <- data_result$data
y_train <- data_result$labels

cat("Generated training data shape:", nrow(trainset), "x", ncol(trainset), "\n")
```

### Compute dissimilarity matrix

```{r compute_dissimilarity}
train_tensor <- torch_tensor(trainset, dtype = torch_float(), device = device)
Dtrain <- torch_cdist(train_tensor, train_tensor, p = 2)

cat("Dissimilarity matrix shape:", Dtrain$shape[1], "x", Dtrain$shape[2], "\n")
```

### Create and train FUSE model

```{r train_model}
model <- FUSE(input_dim = 2, hidden = c(32, 32), dropout = 0.0)
model$to(device)

cat("Training FUSE model...\n")
hist_depth <- model$fit(
  X = train_tensor,
  dissimilarity_matrix = Dtrain,
  partitioned = TRUE,
  allPairs = FALSE,
  num_epochs = 20,
  dsm_resampling = 8,
  anchors_per_pair_global = 64,
  lr = 1e-3,
  device = device,
  quiet = FALSE,
  seed = 379
)
```

### Visualize with contour plots

```{r contour_plots, fig.width=16, fig.height=3}
cat("Generating contour plots...\n")

plots <- plot_homotopy_contours(
  model, trainset, 
  t_values = c(0.0, 0.25, 0.5, 0.75, 1.0),
  grid_n = 200,
  point_size = 1
)

if (is.list(plots)) {
  for (p in plots) {
    print(p)
  }
} else {
  print(plots)
}
```

## Summary

The above example demonstrates FUSE on a synthetic 2D mixture of Gaussian distributions:

**Dataset:** 
- 5000 samples from 4 Gaussian components
- 2D feature space
- Euclidean distance metric

**Model Configuration:**
- Shared encoder: 2 → 32 → 32 → 2
- Global head: learns centrality via anchor-based comparisons
- Local head: learns density via denoising score matching
- Training: 20 epochs with Adam optimizer

**Key Parameters:**
- `partitioned=TRUE`: Deterministic triplet splitting
- `allPairs=FALSE`: 1-to-1 pair matching (vs. Cartesian product)
- `anchors_per_pair_global=64`: Anchor sampling rate
- `dsm_sigma=1.0`: Noise scale for denoising score matching
- `dsm_resampling=8`: Multiple noise resamples per epoch

**Homotopy Interpolation Parameter (t):**
- **t = 0**: Pure global centrality (distribution-level structure)
- **t = 1**: Pure local density (neighborhood-level structure)  
- **0 < t < 1**: Hybrid centrality combining both scales

The contour plots show how the learned centrality measure smoothly transitions from global to local perspectives as t increases, enabling **multi-scale analysis** of data centrality.